---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# kenyaFoodPrices

<!-- badges: start -->
[![Monthly data refresh](https://github.com/m-mburu/kenyaFoodPrices/actions/workflows/dowload_data.yaml/badge.svg)](https://github.com/m-mburu/kenyaFoodPrices/actions/workflows/dowload_data.yaml)
<!-- badges: end -->

This package visualizes food prices and climate conditions in Kenya. Food-price data is sourced from the World Food Programme (WFP) through the [Humanitarian Data Exchange](https://data.humdata.org/dataset/wfp-food-prices-for-kenya?). Rainfall and vegetation indicators originate from WFP and use the World Bank [Joint Food Security Monitor](https://microdata.worldbank.org/catalog/8115) harmonization to current Kenyan administrative boundaries. The app is hosted on shinyapps.io and can be accessed [here](https://mmburu.shinyapps.io/kenyaFoodPrices/).

## Price estimates and coverage

The dashboard defaults to a balanced median: it takes each market's monthly
median, then each county's monthly median, then the national monthly median.
This prevents locations with more raw records from receiving extra weight.
Users can choose a clearly labelled record-weighted mean when that is the
appropriate question. Every price panel reports its calculation and coverage.
Month-on-month changes are shown only when the immediately preceding calendar
month has an observation; gaps are reported as not available rather than
silently treated as one-month changes.

The Climate tab provides county rainfall and vegetation choropleths first, followed by county trends and exploratory lagged relationships with the selected commodity price. See [CLIMATE_IMPLEMENTATION.md](CLIMATE_IMPLEMENTATION.md) for the download, processing, data structure, visual design and limitations.

## Latest climate conditions

The maps below show each county's latest rainfall and vegetation conditions relative to its historical pattern. Negative values indicate below-normal conditions, while positive values indicate above-normal conditions. Open the app's Climate tab for physical values, county trends and commodity-price comparisons.

```{r latest-climate-choropleth, echo=FALSE, fig.width=10, fig.height=5.8, warning=FALSE, message=FALSE}
library(ggplot2)

data(kenya_climate)
data(kenya_counties)

climate_monthly <- data.table::as.data.table(kenya_climate$county_monthly)
climate_lookup <- data.table::as.data.table(kenya_climate$county_lookup)
latest_climate_month <- max(climate_monthly$date, na.rm = TRUE)
latest_climate <- climate_monthly[
  date == latest_climate_month,
  list(adm1_pcode, rainfall_z, ndvi_z)
]
latest_climate <- merge(
  latest_climate,
  climate_lookup[, list(adm1_pcode, county)],
  by = "adm1_pcode",
  all.x = TRUE
)

county_key <- function(x) {
  gsub("[^a-z0-9]", "", tolower(x))
}

climate_shapes <- sf::st_as_sf(kenya_counties)
climate_shapes$county_key <- county_key(climate_shapes$county)
latest_climate$county_key <- county_key(latest_climate$county)
climate_shapes <- merge(
  climate_shapes,
  latest_climate[, list(county_key, rainfall_z, ndvi_z)],
  by = "county_key",
  all.x = TRUE,
  sort = FALSE
)

rainfall_map <- climate_shapes
rainfall_map$indicator <- "Rainfall"
rainfall_map$condition <- rainfall_map$rainfall_z

vegetation_map <- climate_shapes
vegetation_map$indicator <- "Vegetation (NDVI)"
vegetation_map$condition <- vegetation_map$ndvi_z

climate_map <- rbind(
  rainfall_map[, c("indicator", "condition", "geometry")],
  vegetation_map[, c("indicator", "condition", "geometry")]
)

ggplot(climate_map) +
  geom_sf(aes(fill = condition), colour = "white", linewidth = 0.15) +
  facet_wrap(~indicator, nrow = 1) +
  scale_fill_gradient2(
    low = "#B2182B",
    mid = "#F7F7F7",
    high = "#1B7837",
    midpoint = 0,
    name = "Standardised\ncondition"
  ) +
  labs(
    title = paste(
      "Kenya climate conditions —",
      format(latest_climate_month, "%B %Y")
    ),
    subtitle = "Compared with each area's historical conditions",
    caption = "Source: WFP climate indicators, harmonised by the World Bank Joint Food Security Monitor"
  ) +
  coord_sf(datum = NA) +
  theme_void() +
  theme(
    plot.title = element_text(face = "bold"),
    plot.subtitle = element_text(colour = "#555555"),
    strip.text = element_text(face = "bold", size = 11),
    legend.position = "bottom"
  )
```

## Installation

You can install the development version of kenyaFoodPrices from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
#devtools::install_github("m-mburu/kenyaFoodPrices")
```


```{r, fig.width=7, fig.height=4.5, warning=FALSE, message=FALSE}
library(kenyaFoodPrices)
library(ggplot2)
library(dplyr)
library(data.table)
library(ggthemes)
data(ke_food_prices)


ke_food_prices_maize <- ke_food_prices[commodity == "Maize" & pricetype == "Retail" & unit =="KG",]

ke_food_prices_maize[, .(mean_price = mean(price)), by = .(year_quarter_date)] %>%
  ggplot(aes(x = year_quarter_date, y = mean_price)) +
  geom_line(, color = "#66A61E") +
  labs(title = "1KG of Maize Price in Kenya",
       x = "Date",
       y = "Mean Price") +
  scale_x_date(date_labels = "%m/%y", breaks = "24 month") +
 theme_hc()+
  theme(legend.position = "bottom")


```


```{r}
# Example usage of the function
display_time_in_timezone("Africa/Nairobi")


```

## Acknowledgements

- **World Food Programme (WFP)** for publishing Kenya food-price, rainfall and vegetation indicators.
- **World Bank Joint Food Security Monitor team** for harmonising the climate indicators to current Kenyan administrative boundaries.
- **World Health Organization (WHO)** for its nutrition and public-health guidance, which helps frame the food-affordability context of this work.

