---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# kenyaFoodPrices

<!-- badges: start -->
[![R-CMD-check](https://github.com/m-mburu/kenyaFoodPrices/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/m-mburu/kenyaFoodPrices/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

This is a package to visualize food prices in Kenya. The data is sourced from the World Food Programme (WFP) and is available on the Humanitarian Data Exchange (HDX) platform. The app is hosted on shinyapps.io and can be accessed [here](https://mmburu.shinyapps.io/kenyaFoodPrices/).

## Installation

You can install the development version of kenyaFoodPrices from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
#devtools::install_github("m-mburu/kenyaFoodPrices")
```


```{r, fig.width=7, fig.height=4.5, warning=FALSE, message=FALSE}
library(kenyaFoodPrices)
library(ggplot2)
library(dplyr)
library(data.table)
library(ggthemes)
data(ke_food_prices)


ke_food_prices_maize <- ke_food_prices[commodity == "Maize (white)" & pricetype == "Retail",]

ke_food_prices_maize[, .(mean_price = mean(price)), by = .(date)] %>%
  ggplot(aes(x = date, y = mean_price)) +
  geom_line(, color = "steelblue") +
  labs(title = "1KG of Maize Price in Kenya",
       x = "Date",
       y = "Mean Price") +
  scale_x_date(date_labels = "%m/%y", breaks = "24 month") +
 theme_hc()+
  theme(legend.position = "bottom")


```


```{r}
# Example usage of the function
display_time_in_timezone("Africa/Nairobi")


```

- **Thanks to WFP for providing the data on Humanitarian Data Exchange (HDX)**

